Why is it that statistical tests for residuals are not widely used? An explanation using visual inference.

### 2023-01-22 paper_comparison/paper_comparison/paper_comparison.Rmd
remove the use of x_raw

Visual patterns of non-linearity are simulated using four different orders of probabilist's Hermite polynomials ($j = 2, 3, 6, 18$) and four different distributions of $X_{raw}$: (1) $U(-1, 1)$, (2) $N(0, 0.3^2)$, (3) $lognormal(0, 0.6^2)/3$ and (4) $u\{1, 5\}$. A summary of the factors is given in Table \ref{tab:model-factor-table}.

```{r results='asis'}
data.frame(j = c("2", "3", "6", "18", ""),
           x_dist = c("$U(-1, 1)$", "$N(0, 0.3^2)$", "$lognormal(0, 0.6^2)/3$", "$U\\{1, 5\\}$", ""),
           sigma = c("0.25", "1.00", "2.00", "4.00", ""),
           a = c("-1", "0", "1", "", ""),
           b = c("0.25", "1.00", "4.00", "16.00", "64.00"),
           n = c("50", "100", "300", "", "")) %>%
  knitr::kable(col.names = NULL,
               align = "rcrrrr",
               escape = FALSE,
               format = "latex",
               booktabs = TRUE, 
               caption  = 'Description of all factors involved in the non-linear and heteroskedasticity studies.', 
               label = "model-factor-table" ) %>%
  kableExtra::kable_styling(latex_options = "scale_down") %>%
  kableExtra::add_header_above(c("Poly Order (sym1)", #"Order of Hermite polynomial\n(sym1)", 
                                 "Distribution of sym2\n", 
                                 "SD (sym3)", #"Standard deviation of polynomial model\n(sym3)",
                                 "Heteroskedasticity Shape (sym4)", # "Shape of heteroskedasticity\n(sym4)", 
                                 "Heteroskedasticity (sym5)", # Variance factor of heteroskedasticity model\n(sym5)",
                                 "Size (sym6)")) %>% #"Sample size\n(sym6)")) %>%
  as.character() %>%
  gsub("sym1", "$j$", .) %>%
  gsub("sym2", "$X_{raw}$", .) %>%
  gsub("sym3", "$\\\\sigma$", .) %>%
  gsub("sym4", "$a$", .) %>%
  gsub("sym5", "$b$", .) %>%
  gsub("sym6", "$n$", .) %>%
  cat()
```

@@

Different distributions of $X_{raw}$ help enriching the pool of visual patterns as illustrated in Figure \ref{fig:different-dist}. The uniform and the normal distribution are symmetric and commonly assumed in statistical models. The adjusted log-normal distribution provides skewed density, while the discrete uniform distribution provides discreteness in residual plot.

@@

```{r different-dist, fig.height = 2, fig.cap="Variations in fitted values ($X_{raw}$), that might affect perception of residual plots. Four different distribution are used."}

@@

Figure \ref{fig:example-poly-lineup} demonstrates one of the lineups used in experiment I. This lineup is produced by the non-linearity model under $j = 6$ and $X_{raw} \sim N(0.0.3^2)$. The data plot location is $2^3 - 4$. All five subjects correctly identify the data plot from this lineup.

@@

All lineups consist of only null plots are planned to be evaluated by 20 subjects. 
However, presenting only these lineups to subjects are considered to be bad practices as subjects will lose interest quickly. 
Therefore, we plan to collect 6 more evaluations on the 279 lineups with $X_{raw} \sim U(-1,1)$, result in $(36 \times 20 + (4 \times 4 \times 3 + 3 \times 5 \times 3) \times 3 \times 6) / (20-2) = `r (36 * 20 + (4 * 4 * 3 + 3 * 5 * 3) * 3 * 6) / 18`$ subjects recruited for experiment III.

@@

There are `r 160 * 18`, `r 150 * 18` and `133 * 18` lineups evaluation made by 160, 150 and 133 subjects recruited for experiment I, II and III respectively. In the total of `r nrow(filter(vi_survey, !attention_check))` lineup evaluations, `r nrow(filter(vi_survey, type == "polynomial", !attention_check))` use lineups produced by the non-linearity model, `r nrow(filter(vi_survey, type == "heteroskedasticity", !attention_check))` use lineups produced by the heteroskedasticity model, and `r nrow(filter(vi_survey, null_lineup, !attention_check))` use null lineups. Besides, there are `r nrow(filter(vi_survey, attention_check))` attention checks not included in the following analysis. The collated dataset is provided in `vi_survey` of the `visage` `R` package.

In the following analysis, lineups with $X_{raw} \sim U(-1,1)$ will be the focus of next two sections, because visual patterns are more likely to be revealed with uniform distribution. Besides, six more evaluations are collected for the uniform distribution in experiment III, which will produce more reliable and stable results. Analysis of other distributions can be found in Section \ref{effect-of-the-distribution-of-the-regressor}.

@@

Figure \ref{fig:polypower} shows the estimated power of visual test on lineups produced by the non-linearity model with $X_{raw} \sim U(-1,1)$, against the natural logarithm of the effect $log_e(E)$, with a 5% significance level. At the bottom of the figure \ref{fig:polypower}, there are a sequence of example residual plots with increasing levels of $log_e(E)$. Readers can evaluate them from left to right and determine at which level the departure from a good residual plot becomes detectable.

>>>>

Visual patterns of non-linearity are simulated using four different orders of probabilist's Hermite polynomials ($j = 2, 3, 6, 18$). Four distributions of fitted values are also considered by using $X_{raw}$ follows different distributions: (1) $U(-1, 1)$, (2) $N(0, 0.3^2)$, (3) $lognormal(0, 0.6^2)/3$ and (4) $u\{1, 5\}$. A summary of the factors is given in Table \ref{tab:model-factor-table}.

```{r results='asis'}
data.frame(j = c("2", "3", "6", "18", ""),
           x_dist = c("Uniform", "Normal", "Skewed", "Discrete uniform", ""),
           sigma = c("0.25", "1.00", "2.00", "4.00", ""),
           a = c("-1", "0", "1", "", ""),
           b = c("0.25", "1.00", "4.00", "16.00", "64.00"),
           n = c("50", "100", "300", "", "")) %>%
  knitr::kable(col.names = NULL,
               align = "rcrrrr",
               escape = FALSE,
               format = "latex",
               booktabs = TRUE, 
               caption  = 'Description of all factors involved in the non-linear and heteroskedasticity studies.', 
               label = "model-factor-table" ) %>%
  kableExtra::kable_styling(latex_options = "scale_down") %>%
  kableExtra::add_header_above(c("Poly Order (sym1)", #"Order of Hermite polynomial\n(sym1)", 
                                 "Distribution of fitted values", 
                                 "SD (sym3)", #"Standard deviation of polynomial model\n(sym3)",
                                 "Heteroskedasticity Shape (sym4)", # "Shape of heteroskedasticity\n(sym4)", 
                                 "Heteroskedasticity (sym5)", # Variance factor of heteroskedasticity model\n(sym5)",
                                 "Size (sym6)")) %>% #"Sample size\n(sym6)")) %>%
  as.character() %>%
  gsub("sym1", "$j$", .) %>%
  gsub("sym3", "$\\\\sigma$", .) %>%
  gsub("sym4", "$a$", .) %>%
  gsub("sym5", "$b$", .) %>%
  gsub("sym6", "$n$", .) %>%
  cat()
```

@@

Different distributions of fitted values help enriching the pool of visual patterns as illustrated in Figure \ref{fig:different-dist}. The uniform and the normal distribution are symmetric and commonly assumed in statistical models. The adjusted log-normal distribution provides skewed density, while the discrete uniform distribution provides discreteness in residual plot.

@@

```{r different-dist, fig.height = 2, fig.cap="Variations in fitted values, that might affect perception of residual plots. Four different distribution are used."}

@@

Figure \ref{fig:example-poly-lineup} demonstrates one of the lineups used in experiment I. This lineup is produced by the non-linearity model under $j = 6$ and fitted values follow a uniform distribution. The data plot location is $2^3 - 4$. All five subjects correctly identify the data plot from this lineup.

@@

All lineups consist of only null plots are planned to be evaluated by 20 subjects. 
However, presenting only these lineups to subjects are considered to be bad practices as subjects will lose interest quickly. 
Therefore, we plan to collect 6 more evaluations on the 279 lineups with fitted values follow a uniform distribution, result in $(36 \times 20 + (4 \times 4 \times 3 + 3 \times 5 \times 3) \times 3 \times 6) / (20-2) = `r (36 * 20 + (4 * 4 * 3 + 3 * 5 * 3) * 3 * 6) / 18`$ subjects recruited for experiment III.

@@

There are `r 160 * 18`, `r 150 * 18` and `r 133 * 18` lineups evaluation made by 160, 150 and 133 subjects recruited for experiment I, II and III respectively. In the total of `r nrow(filter(vi_survey, !attention_check))` lineup evaluations, `r nrow(filter(vi_survey, type == "polynomial", !attention_check))` use lineups produced by the non-linearity model, `r nrow(filter(vi_survey, type == "heteroskedasticity", !attention_check))` use lineups produced by the heteroskedasticity model, and `r nrow(filter(vi_survey, null_lineup, !attention_check))` use null lineups. Besides, there are `r nrow(filter(vi_survey, attention_check))` attention checks not included in the analysis. The collated dataset is provided in `vi_survey` of the `visage` `R` package.

In the following analysis, lineups with fitted values follows a uniform distribution will be the focus of next two sections. Visual patterns are more likely to be revealed under a uniform distribution. Additionally, we have collected extra evaluations on these lineups, which will result in more reliable analysis. Analysis of lineups with other fitted value distributions can be found in Section \ref{fitted-value-distribution}.

@@

Figure \ref{fig:polypower} shows the estimated power of visual test on lineups produced by the non-linearity model with fitted values follow a uniform distribution, against the natural logarithm of the effect $log_e(E)$, with a 5% significance level. At the bottom of the figure \ref{fig:polypower}, there are a sequence of example residual plots with increasing levels of $log_e(E)$. Readers can evaluate them from left to right and determine at which level the departure from a good residual plot becomes detectable.

### 2023-01-22 paper_comparison/paper_comparison/paper_comparison.tex
modify plot for different b
remove the use of x_raw

Visual patterns of non-linearity are simulated using four different
orders of probabilist's Hermite polynomials (\(j = 2, 3, 6, 18\)) and
four different distributions of \(X_{raw}\): (1) \(U(-1, 1)\), (2)
\(N(0, 0.3^2)\), (3) \(lognormal(0, 0.6^2)/3\) and (4) \(u\{1, 5\}\). A
summary of the factors is given in Table \ref{tab:model-factor-table}.

@@

\caption{\label{tab:model-factor-table}Description of all factors involved in the non-linear and heteroskedasticity studies.}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{rcrrrr}
\toprule
\multicolumn{1}{c}{Poly Order ($j$)} & \multicolumn{1}{c}{\makecell[c]{Distribution of $X_{raw}$\\}} & \multicolumn{1}{c}{SD ($\sigma$)} & \multicolumn{1}{c}{Heteroskedasticity Shape ($a$)} & \multicolumn{1}{c}{Heteroskedasticity ($b$)} & \multicolumn{1}{c}{Size ($n$)} \\
\cmidrule(l{3pt}r{3pt}){1-1} \cmidrule(l{3pt}r{3pt}){2-2} \cmidrule(l{3pt}r{3pt}){3-3} \cmidrule(l{3pt}r{3pt}){4-4} \cmidrule(l{3pt}r{3pt}){5-5} \cmidrule(l{3pt}r{3pt}){6-6}
2 & $U(-1, 1)$ & 0.25 & -1 & 0.25 & 50\\
3 & $N(0, 0.3^2)$ & 1.00 & 0 & 1.00 & 100\\
6 & $lognormal(0, 0.6^2)/3$ & 2.00 & 1 & 4.00 & 300\\
18 & $U\{1, 5\}$ & 4.00 &  & 16.00 & \\
 &  &  &  & 64.00 & \\
\bottomrule
\end{tabular}}
\end{table}

@@

Different distributions of \(X_{raw}\) help enriching the pool of visual
patterns as illustrated in Figure \ref{fig:different-dist}. The uniform
and the normal distribution are symmetric and commonly assumed in
statistical models. The adjusted log-normal distribution provides skewed
density, while the discrete uniform distribution provides discreteness
in residual plot.

@@

\caption{Variations in fitted values ($X_{raw}$), that might affect perception of residual plots. Four different distribution are used.}\label{fig:different-dist}
\end{figure}

Figure \ref{fig:example-poly-lineup} demonstrates one of the lineups
used in experiment I. This lineup is produced by the non-linearity model
under \(j = 6\) and \(X_{raw} \sim N(0.0.3^2)\). The data plot location
is \(2^3 - 4\). All five subjects correctly identify the data plot from
this lineup.

@@

All lineups consist of only null plots are planned to be evaluated by 20
subjects. However, presenting only these lineups to subjects are
considered to be bad practices as subjects will lose interest quickly.
Therefore, we plan to collect 6 more evaluations on the 279 lineups with
\(X_{raw} \sim U(-1,1)\), result in
\((36 \times 20 + (4 \times 4 \times 3 + 3 \times 5 \times 3) \times 3 \times 6) / (20-2) = 133\)
subjects recruited for experiment III.

@@

There are 2880, 2700 and \texttt{133\ *\ 18} lineups evaluation made by
160, 150 and 133 subjects recruited for experiment I, II and III
respectively. In the total of 7974 lineup evaluations, 3744 use lineups
produced by the non-linearity model, 4230 use lineups produced by the
heteroskedasticity model, and 720 use null lineups. Besides, there are
886 attention checks not included in the following analysis. The
collated dataset is provided in \texttt{vi\_survey} of the
\texttt{visage} \texttt{R} package.

In the following analysis, lineups with \(X_{raw} \sim U(-1,1)\) will be
the focus of next two sections, because visual patterns are more likely
to be revealed with uniform distribution. Besides, six more evaluations
are collected for the uniform distribution in experiment III, which will
produce more reliable and stable results. Analysis of other
distributions can be found in Section
\ref{effect-of-the-distribution-of-the-regressor}.

@@

Figure \ref{fig:polypower} shows the estimated power of visual test on
lineups produced by the non-linearity model with
\(X_{raw} \sim U(-1,1)\), against the natural logarithm of the effect
\(log_e(E)\), with a 5\% significance level. At the bottom of the figure
\ref{fig:polypower}, there are a sequence of example residual plots with
increasing levels of \(log_e(E)\). Readers can evaluate them from left
to right and determine at which level the departure from a good residual
plot becomes detectable.

>>>>

Visual patterns of non-linearity are simulated using four different
orders of probabilist's Hermite polynomials (\(j = 2, 3, 6, 18\)). Four
distributions of fitted values are also considered by using \(X_{raw}\)
follows different distributions: (1) \(U(-1, 1)\), (2) \(N(0, 0.3^2)\),
(3) \(lognormal(0, 0.6^2)/3\) and (4) \(u\{1, 5\}\). A summary of the
factors is given in Table \ref{tab:model-factor-table}.

@@

\caption{\label{tab:model-factor-table}Description of all factors involved in the non-linear and heteroskedasticity studies.}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{rcrrrr}
\toprule
\multicolumn{1}{c}{Poly Order ($j$)} & \multicolumn{1}{c}{Distribution of fitted values} & \multicolumn{1}{c}{SD ($\sigma$)} & \multicolumn{1}{c}{Heteroskedasticity Shape ($a$)} & \multicolumn{1}{c}{Heteroskedasticity ($b$)} & \multicolumn{1}{c}{Size ($n$)} \\
\cmidrule(l{3pt}r{3pt}){1-1} \cmidrule(l{3pt}r{3pt}){2-2} \cmidrule(l{3pt}r{3pt}){3-3} \cmidrule(l{3pt}r{3pt}){4-4} \cmidrule(l{3pt}r{3pt}){5-5} \cmidrule(l{3pt}r{3pt}){6-6}
2 & Uniform & 0.25 & -1 & 0.25 & 50\\
3 & Normal & 1.00 & 0 & 1.00 & 100\\
6 & Skewed & 2.00 & 1 & 4.00 & 300\\
18 & Discrete uniform & 4.00 &  & 16.00 & \\
 &  &  &  & 64.00 & \\
\bottomrule
\end{tabular}}
\end{table}

@@

Different distributions of fitted values help enriching the pool of
visual patterns as illustrated in Figure \ref{fig:different-dist}. The
uniform and the normal distribution are symmetric and commonly assumed
in statistical models. The adjusted log-normal distribution provides
skewed density, while the discrete uniform distribution provides
discreteness in residual plot.

@@

\caption{Variations in fitted values, that might affect perception of residual plots. Four different distribution are used.}\label{fig:different-dist}
\end{figure}

Figure \ref{fig:example-poly-lineup} demonstrates one of the lineups
used in experiment I. This lineup is produced by the non-linearity model
under \(j = 6\) and fitted values follow a uniform distribution. The
data plot location is \(2^3 - 4\). All five subjects correctly identify
the data plot from this lineup.

@@

All lineups consist of only null plots are planned to be evaluated by 20
subjects. However, presenting only these lineups to subjects are
considered to be bad practices as subjects will lose interest quickly.
Therefore, we plan to collect 6 more evaluations on the 279 lineups with
fitted values follow a uniform distribution, result in
\((36 \times 20 + (4 \times 4 \times 3 + 3 \times 5 \times 3) \times 3 \times 6) / (20-2) = 133\)
subjects recruited for experiment III.

@@

There are 2880, 2700 and 2394 lineups evaluation made by 160, 150 and
133 subjects recruited for experiment I, II and III respectively. In the
total of 7974 lineup evaluations, 3744 use lineups produced by the
non-linearity model, 4230 use lineups produced by the heteroskedasticity
model, and 720 use null lineups. Besides, there are 886 attention checks
not included in the analysis. The collated dataset is provided in
\texttt{vi\_survey} of the \texttt{visage} \texttt{R} package.

In the following analysis, lineups with fitted values follows a uniform
distribution will be the focus of next two sections. Visual patterns are
more likely to be revealed under a uniform distribution. Additionally,
we have collected extra evaluations on these lineups, which will result
in more reliable analysis. Analysis of lineups with other fitted value
distributions can be found in Section \ref{fitted-value-distribution}.

@@

Figure \ref{fig:polypower} shows the estimated power of visual test on
lineups produced by the non-linearity model with fitted values follow a
uniform distribution, against the natural logarithm of the effect
\(log_e(E)\), with a 5\% significance level. At the bottom of the figure
\ref{fig:polypower}, there are a sequence of example residual plots with
increasing levels of \(log_e(E)\). Readers can evaluate them from left
to right and determine at which level the departure from a good residual
plot becomes detectable.
